Chord: A Versatile Platform for Program Analysis

### 2011-04-15 main/doc/arch.tex
more documentation

\subsubsection*{Chord Properties}

@@

\subsubsection*{Java Program Representation}

@@

\subsubsection*{Analysis Scope Construction}

@@

\subsubsection*{Writing and Running Analyses}

@@

Chord can be invoked in one of two modes: {\it classic} or {\it modern}.  These
two modes defer in the semantics of dependencies between analyses.  In
particular, the classic mode is simpler to understand for novice users (the
dependencies are only data dependencies) but has a sequential runtime, whereas
the modern mode is harder to understand (there are both data and control
dependencies) but has a parallel runtime that is capable of running analyses
without dependencies between them in parallel.  The parallel runtime is based on
\xlink{Habanero-Java}{http://habanero.rice.edu/hj.html}, and the semantics of
the dependencies between analyses is based on the
\xlink{Habanero Concurrent Collections (CnC)}{http://habanero.rice.edu/cnc.html}
declarative parallel programming model.

Chapter \ref{chap:project} expands upon the modular architecture of analyses in
Chord that is described above.

Chord provides various {\it analysis templates}: classes containing boilerplate
code that can be extended by users to rapidly prototype different kinds of
analyses.  An example is class RHSAnalysis, named after [Reps, Horowitz, and
Sagiv 1995], which can be extended by users to write a summary-based
inter-procedural context-sensitive static analysis by merely specifying the
abstract domain and intra-procedural transfer functions.  Another example is
DynamicAnalysis, which can be extended by users to write a dynamic analysis by
merely specifying which of various provided events to instrument, and the
transfer functions for those events.

Chapters \ref{chap:writing} and \ref{chap:running} describe how to write and run
your own analyses in Chord using the provided analysis templates.

\subsubsection*{Dynamic Analysis}

@@

\subsubsection*{Datalog Analysis}

>>>>

{\bf Chord Properties}

@@

{\bf Java Program Representation}

@@

{\bf Analysis Scope Construction}

@@

{\bf Writing and Running Analyses}

@@

Chord can be invoked in one of two modes: {\it classic} or {\it modern}.  These
two modes defer in the semantics of dependencies between analyses.  In
particular, the classic mode is simpler to understand for novice users (the
dependencies are only data dependencies) but has a sequential runtime, whereas
the modern mode is harder to understand (there are both data and control
dependencies) but has a parallel runtime that is capable of running analyses
without dependencies between them in parallel.  The parallel runtime is based on
\xlink{Habanero-Java}{http://habanero.rice.edu/hj.html}, and the semantics of
the dependencies between analyses is based on the
\xlink{Habanero Concurrent Collections (CnC)}{http://habanero.rice.edu/cnc.html}
declarative parallel programming model.  Chapter \ref{chap:project} expands upon
the modular architecture of analyses in Chord.

Chord provides various {\it analysis templates}: classes containing boilerplate
code that can be extended by users to rapidly prototype different kinds of
analyses.  An example is class RHSAnalysis, named after [Reps, Horowitz, and
Sagiv 1995], which can be extended by users to write a summary-based
inter-procedural context-sensitive static analysis by merely specifying the
abstract domain and intra-procedural transfer functions.  Another example is
DynamicAnalysis, which can be extended by users to write a dynamic analysis by
merely specifying which of various provided events to instrument, and the
transfer functions for those events.  Chapters \ref{chap:writing} and
\ref{chap:running} describe how to write and run your own analyses in Chord
using the provided analysis templates.

{\bf Dynamic Analysis}

@@

{\bf Datalog Analysis}

### 2011-04-15 main/doc/predefined.tex
more documentation

\xname{predefined_analyses}
\chapter{Predefined Analyses}
\label{chap:predefined-analyses}

This chapter describes various predefined analyses in Chord.

@@

Chord offers several choices for computing points-to and call-graph information of Java programs.
In each of these choices, points-to and call-graph information is computed simultaneously
(called ``on-the-fly call-graph construction" in the literature in contrast to ``ahead-of-time call-graph
construction" in which the call graph is computed first followed by points-to information).
On-the-fly approaches are more precise because, in a dynamically dispatching language like Java,
as more points-to facts are discovered, more (dynamically dispatched) methods are deemed reachable,
thereby growing the call graph; the code in these newly added methods in turn results in more
points-to facts.

@@

\begin{itemize}
\item
{\it Call-graph information:}
\begin{itemize}
\item rootM subset M contains the set of entry methods that may be reachable; this includes the program's main method
as well as each static initializer method that may be reachable from the main
method.
\item reachableM subset M contains the set of methods that may be reachable from the program's main method.
\item IM subset (I $\times$ M) contains tuples \texttt{(i,m)} such that call site \texttt{i} may call method \texttt{m}.
%\item MM subset (M $\times$ M) contains tuples \texttt{(m1,m2)} such that method \texttt{m1} may call method \texttt{m2}.
%\item reachableI subset I contains the set of call sites that may be reachable.
%\item reachableT subset T contains the set of classes that may be reachable.
\end{itemize}
\item
{\it Points-to information:}
\begin{itemize}
\item FH subset (F $\times$ H) contains tuples \texttt{(f,h)} such that static field (i.e. global variable) \texttt{f} may point to an object allocated at site \texttt{h}.
\item VH subset (V $\times$ H) contains tuples \texttt{(v,h)} such that local variable \texttt{v} may point to an object allocated at site \texttt{h}.
\item HFH subset (H $\times$ F $\times$ H) contains tuples \texttt{(h1,f,h2)} such that instance field \texttt{f} of some object allocated at
site \texttt{h1} may point to some object allocated at site \texttt{h2}.
\end{itemize}
\end{itemize}

>>>>

\xname{predefined}
\chapter{Predefined Analyses}
\label{chap:predefined}

Chord provides many standard analyses.
This chapter first explains how to run any such analysis and then provides
descriptions of various predefined analyses.

\section{Running an Analysis}
\label{sec:running-predefined}

Each predefined analysis in Chord has a unique name that can be used to run the
analysis from the command-line.  The following command runs the analysis named
\code{<ANALYSIS_NAME>} on the program specified by directory \code{<WORK_DIR>}
(see Chapter \ref{chap:setup} for how to setup a program):

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=<ANALYSIS_NAME> run
\end{verbatim}
\end{framed}

For instance, the following command runs a basic may-alias and call-graph
analysis (called 0CFA) provided in Chord:

This instructs Chord to run the analysis named \code{cipa-0cfa-dlog},
which is defined in file \code{main/src/chord/analyses/alias/cipa_0cfa.dlog}.

The output of the above command is of the form:

\begin{framed}
{\small
\begin{verbatim}
Buildfile: build.xml

run:
     [java] Chord run initiated at: Mar 13, 2011 10:31:08 PM
     [java] ENTER: cipa-0cfa-dlog
     [java] ENTER: T
     [java] ENTER: RTA
     [java] Iteration: 0
     [java] Iteration: 1
     [java] Iteration: 2
     [java] LEAVE: RTA
     [java] SAVING dom T size: 1386
     [java] LEAVE: T
     [java] ENTER: F
     [java] SAVING dom F size: 4120
     [java] LEAVE: F
     ...
     [java] ENTER: MputStatFldInst
     [java] SAVING rel MputStatFldInst size: 739
     [java] LEAVE: MputStatFldInst
     [java] ENTER: statIM
     [java] SAVING rel statIM size: 3359
     [java] LEAVE: statIM
     [java] Starting command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
     [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
     [java] Relation FH: 137 nodes, 8.0 elements (H0,F0)
     [java] Relation HFH: 199 nodes, 35.0 elements (H0,F0,H1)
     [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
     [java] Finished command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
     [java] LEAVE: cipa-0cfa-dlog
     [java] Chord run completed at: Mar 13, 2011 10:31:36 PM
     [java] Total time: 00:00:27:671 hh:mm:ss:ms

BUILD SUCCESSFUL
Total time: 28 seconds
\end{verbatim}
}
\end{framed}

Each analysis in Chord is written modularly, independent of other analyses,
along with lightweight annotations specifying the inputs and outputs of the
analysis.  Chord's runtime automatically computes producer-consumer
relationships between analyses (e.g., determines which analysis produces as
output a result that is needed as input by another analysis).  Before running a
desired analysis (such as 0CFA in the above example), Chord recursively runs
other analyses until the inputs to the desired analysis have been computed; it
finally runs the desired analysis to produce the outputs of that analysis.

The 0CFA analysis consumes and produces multiple {\it program relations}.  The
consumed program relations include \code{MputStatFldInst} and \code{statIM},
each of which is produced by a separate imperative analysis with the
corresponding name, and the produced program relations include \code{VH},
\code{FH}, \code{HFH}, and \code{IM}.  We next briefly discuss these relations.

The program relations consumed by the 0CFA analysis contain basic program facts.
For instance, \code{MputStatFldInst} is a relation containing each tuple
($m$,$f$,$v$) such that method $m$ in the input Java program contains a
putstatic instruction of the form ``$f$ = $v$", while \code{statIM} is a
relation containing each tuple ($i$,$m$) such that $m$ is the target method of
invokestatic instruction $i$.

The program relations produced by the 0CFA analysis represent points-to
information and the call graph of the input Java program as computed by the
analysis.  Specifically, relations \code{VH}, \code{FH}, and \code{HFH}
represent points-to information for local variables, static fields, and instance
fields, respectively, while relation \code{IM} represents the call graph,
namely, it contain each tuple ($i$,$m$) such that $m$ is a possible target
method of call site $i$.

Metavariables $m$, $f$, $i$, and $v$ above range over entities in so-called
{\it program domains} \code{M}, \code{F}, \code{I}, and \code{V}, respectively.
A program domain is a set of entities of a certain kind in the input Java
program.  For instance, \code{M} is the domain representing the set of all
methods in the input Java program, \code{F} is the domain representing the set
of all fields, \code{I} is the domain representing the set of all call sites in
methods in \code{M}, and \code{V} is the domain representing the set of all
local variables of reference type in methods in \code{M}.  Each of these domains
is produced by a separate Java analysis with the corresponding name.  Notice
that the analyses producing these domains run upfront because these domains are
consumed by the analyses that produce relations such as \code{MputStatFldInst}
and \code{statIM}, which in turn are consumed by the desired 0CFA analysis.

During execution, Chord dumps intermediate and final results to files in the
directory specified by property \code{chord.out.dir}, whose default value is
\code{[chord.work.dir]/chord_output/} and typically does not need to be changed
by users.  For the above example, this directory is
\code{<WORK_DIR>/chord_output/}.

The verbosity of Chord's output above is controlled by property
\code{chord.verbose}, whose default value is 1.  At verbosity level 0, the above
command produces less voluminous output of the form:

\begin{framed}
{\small
\begin{verbatim}
Buildfile: build.xml

run:
     [java] Chord run initiated at: Mar 13, 2011 10:35:01 PM
     [java] Chord run completed at: Mar 13, 2011 10:35:28 PM
     [java] Total time: 00:00:26:297 hh:mm:ss:ms

BUILD SUCCESSFUL
Total time: 26 seconds
\end{verbatim}
}
\end{framed}

@@

Chord offers several choices for computing points-to and call-graph information
of Java programs.  In each of these choices, the points-to and call-graph
information is computed simultaneously (called ``on-the-fly call-graph
construction" in the literature in contrast to ``ahead-of-time call-graph
construction" in which the call graph is computed first followed by points-to
information).  On-the-fly approaches are more precise because, in a dynamically
dispatching language like Java, as more points-to facts are discovered, more
(dynamically dispatched) methods are deemed reachable, thereby growing the call
graph; the code in these newly added methods in turn results in more points-to
facts.

@@

\begin{itemize}
\item {\it Call-graph information:}
\begin{itemize}
\item
rootM subset M contains the set of entry methods that may be reachable; this
includes the program's main method as well as each static initializer method
that may be reachable from the main method.
\item
reachableM subset M contains the set of methods that may be reachable from the
program's main method.
\item
IM subset (I $\times$ M) contains tuples \texttt{(i,m)} such that call site
\texttt{i} may call method \texttt{m}.
%\item
% MM subset (M $\times$ M) contains tuples \texttt{(m1,m2)} such that
%method \texttt{m1} may call method \texttt{m2}.
%\item
%reachableI subset I contains the set of call sites that may be reachable.
%\item
%reachableT subset T contains the set of classes that may be reachable.
\end{itemize}
\item
{\it Points-to information:}
\begin{itemize}
\item
FH subset (F $\times$ H) contains tuples \texttt{(f,h)} such that static field
(i.e.  global variable) \texttt{f} may point to an object allocated at site
\texttt{h}.
\item
VH subset (V $\times$ H) contains tuples \texttt{(v,h)} such that local variable
\texttt{v} may point to an object allocated at site \texttt{h}.
\item
HFH subset (H $\times$ F $\times$ H) contains tuples \texttt{(h1,f,h2)} such
that instance field \texttt{f} of some object allocated at site \texttt{h1} may
point to some object allocated at site \texttt{h2}.
\end{itemize}
\end{itemize}
